---
title: "Data Analyltics in Finance"
author: "Jeff Bostwick"
date: "11/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Holt Winter's Forecasting Method
 
#### Loading Packages
```{r Loadiing Packages, echo=TRUE, message=FALSE, warning=FALSE}
library(forecast)
library(fma)
library(datasets)
library(ggplot2)
```
 

#### View Sample Dataset
```{r Sample Data, echo=TRUE, message=FALSE, warning=FALSE}
airpass
```
 

#### Plot the Dataset

```{r}
autoplot(airpass)
```
 
 
### Simple Exponential Smoothing Example
#### Creating a varaiable called SES5 that is simple expontntial smoothing.  h accounts for the number of lags.
 
```{r}
ses5 <- ses(airpass, h=5)
ses5
```

Here are some various measures of error.  The RMSE is the one that we should focus on
```{r}
accuracy(ses5)
```
 
 
### Creating a Forecast Plot with Simple Exponential Smoothing

```{r Simple Exponential Smoothing Example, echo=TRUE, message=FALSE, warning=FALSE}
autoplot(ses5) +
autolayer(fitted(ses5), series = "Fitted")
```

From this we can see an upward trend and see that there is a seasonal cycle that has an upward trend
The fitted values are in Red.  The Dark Blue is the 80% confidence interval. 
The light blue is the 95% confidence interval.  The line is the predicted value = 431.99
 
### Now we can try to adujust to 25 lags
```{r}
ses25 <- ses(airpass, h=25)
autoplot(ses25)
```

 
## Holt's Linear Trend Model
#### This will add the linear trend to our model
 
```{r Creating Holt Method Graph}
holt5 <- holt(airpass, h=5) 
autoplot(holt5) + autolayer(fitted(holt5), series = "Fitted")
```
 
#### There is evidence that the Holt Linear Trend method overestimates the predicted values.
#### There is research that says that dampening the trend can help with accuracy.
```{r Damped Values, echo=TRUE, message=FALSE, warning=FALSE}
holt5damped <- holt(airpass, damped = TRUE, phi = 0.9, h=15)
```
 
```{r Plot ith Holt Method and Damped Value, echo=TRUE, message=FALSE, warning=FALSE}
autoplot(airpass) + autolayer(holt5, series="Holt's Method", PI=FALSE)+
autolayer(holt5damped, series="Damped Holt's Method", PI=FALSE) +
ggtitle("Forecasts from Holt's Method") +
guides(colour=guide_legend(title="Forecast"))
```

 
## Holt's Seasonal Trend Model
#### Now we are adding in the seasonalilty to the model.  There are 2 options, additive and multiplicative models.
   
```{r echo=TRUE}
hw1 <- hw(airpass, seasonal = "additive")
hw2 <- hw(airpass, seasonal = "multiplicative")
```
 
#### Plot the Holt's Seasonal Trend Model
```{r Holts Seasonal Trend Model, echo=TRUE}
autoplot(airpass) +
  autolayer(hw1, series = "HW additive forecasts", PI = FALSE) +
  autolayer(hw2, series = "HW multiplicative forecasts", PI = FALSE) +
  guides(colour = guide_legend(title="Forecast"))
```

Just by looking at the graph above, it looks like the multiplicative model is a better fit based off of past data.


